\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics

### 2016-08-16 journalSwarmControl/experiment.tex
fixed bf u
Changed format

 A corresponding reward function gives a high reward to the goal state,  
 a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,u) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to a state with an obstacle} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 $\mathbf{u}$ is the action and $x$ is the current state. 
  Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| \mathbf{u},x)].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration}\\
\text{end}& \nonumber
\end{align}
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

\begin{figure}
\centering
\begin{overpic}[scale=0.19]{GradientView.png}
\end{overpic}
\begin{overpic}[scale=0.19]{PolicyIter.png}
\end{overpic}
\begin{overpic}[scale=0.262]{MDPmap.pdf}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:BFSGradient}The BFS algorithm finds the shortest path for the moveable object  to compute gradient vectors (left). Modeling as an MDP enables encoding penalties for being near obstacles. (Middle) The control policy from value iteration. (Right) The vision algorithm detects obstacles in the hardware setup. This map is used to produce the value function and control policy shown. 
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure*}
\begin{center}
	\begin{overpic}[width=0.6\columnwidth]{stateMachine.pdf}\put(-12,25){A}\end{overpic}\\
	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true
	\put(-6,85){B}
	\end{overpic}
	~~
	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}
	\put(-6,85){C}
	\end{overpic}
\end{center}
\vspace{-0.5em}
\caption{\label{fig:Region}  Outlier rejection state machine and regions.
}
\end{figure*}

@@

\begin{wrapfigure}{R}{0.3\textwidth}
  \vspace{-20pt}
  \begin{center}
\begin{overpic}[width=0.3\columnwidth]{AllShapes.pdf}\end{overpic}
  \end{center}
  \vspace{-1em}
\caption{\label{fig:Shapes} Six equal-area objects were tested. %Six object shapes were tested.
\vspace{-1em}
}
\end{wrapfigure}

@@

\begin{algorithm}
\caption{Object-manipulation controller for a robotic swarm.}\label{alg:BlockPushing}
\begin{algorithmic}[1]
\Require Knowledge of moveable object's center of mass $\mathbf{b}$; swarm mean $[\bar{x},\bar{y}]$ and variance $[\sigma_x^2, \sigma_y^2]$, each calculated using the regions function from \S \ref{subsec:OutlierRejection};  map of the environment
\State Compute optimal policy for object, according to \S \ref{subsec:objectpolicy}
\While{$\mathbf{b}$ is not in goal region}
\State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$
\If {$\sigma^2 > \sigma_{max}^2$}
\While{$\sigma^2 > \sigma_{min}^2$}
\State $ [x_{goal}, y_{goal}] \gets $ the nearest corner in current region
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndWhile
\Else  
%\If {$\mathrm{distance}( \mathbf{b}, [x_{goal}, y_{goal}] ) > k_1 r_b$}
%	\State$r_p \gets k_2 r_b$  \Comment{guarded move}
%	\Else
%	\State$r_p \gets k_3 r_b$  \Comment{pushing move}
%	\EndIf
%\State$r_p \gets k_3 r_b$  \Comment{pushing move}
\State Calculate $\mathbf{D}(\mathbf{b})$  \Comment{direction for object at $\mathbf{b}$}
\State Apply \eqref{eq:potentialfield}   \Comment{potential field to push object in correct direction}
\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

@@

\begin{figure*}
\centering
%\renewcommand{\figwid}{0.19\columnwidth}
%\href{http://youtu.be/tCej-9e6-4o}{\begin{overpic}[width =\figwid]{story1.png}\put(6,15){T = 5 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story2.png}\put(6,15){T = 12 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story3.png}\put(6,15){T = 20 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story4.png}\put(6,15){T = 25 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story5.png}\put(6,15){T = 33 s}
%\end{overpic}}
\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}
\end{overpic}
\vspace{-2em}
\caption{\label{fig:story}\href{http://youtu.be/tCej-9e6-4o}{Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See video in multimedia extension.}
%\vspace{-2em}
}
\end{figure*}

@@

\begin{figure*}
\centering
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =0.45\columnwidth]{SimVeryNum.pdf}\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =0.45\columnwidth]{SimVeryNoise.pdf}\put(1,55){b)}
\end{overpic}
\begin{overpic}[width =0.45\columnwidth]{SimVeryWeight.pdf}\put(1,55){c)}
\end{overpic}
\begin{overpic}[width =0.45\columnwidth]{SimVeryShape.pdf}\put(1,55){d)}

>>>>

 A corresponding reward function gives a high reward to the goal state,  
 a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to a state with an obstacle} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 $\mathbf{u}$ is the action and $x$ is the current state. 
  Value iteration iteratively learns the value $\hat{V}(x)$ for the object being in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| \mathbf{u},x)].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration} \nonumber\\
\text{end}& 
\end{align}
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

\begin{figure}
\centering
\renewcommand{\figwid}{3cm}
\begin{overpic}[height=\figwid]{GradientView.png}
\end{overpic}
\begin{overpic}[height=\figwid]{PolicyIter.png}
\end{overpic}
\begin{overpic}[height=\figwid]{MDPmap.pdf}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left). Modeling as an MDP enables encoding penalties for being near obstacles. (Middle) The control policy from value iteration. (Right) The vision algorithm detects obstacles in the hardware setup. This map is used to produce the value function and control policy shown. 
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure}
\begin{center}
	\begin{overpic}[width=0.9\columnwidth]{stateMachine.pdf}\put(-12,25){A}\end{overpic}\\
	~~\begin{overpic}[width=0.45\columnwidth]{mainRegions.pdf} %,grid=true
	\put(-6,85){B}
	\end{overpic}
	~~
	\begin{overpic}[width=0.45\columnwidth]{transferRegions.pdf}
	\put(-6,85){C}
	\end{overpic}
\end{center}
\vspace{-0.5em}
\caption{\label{fig:Region}  Outlier rejection state machine and regions.
}
\end{figure}

@@

\begin{figure}
%  \vspace{-20pt}
  \begin{center}
\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\end{overpic}
  \end{center}
%  \vspace{-1em}
\caption{\label{fig:Shapes} The six equal-area objects tested in simulation. %Six object shapes were tested.
\vspace{-1em}
}
\end{figure}

@@

\begin{algorithm}
\caption{Object-manipulation controller for a robotic swarm.}\label{alg:BlockPushing}
\begin{algorithmic}[1]
\Require Knowledge of moveable object's center of mass $\mathbf{b}$; swarm mean $[\bar{x},\bar{y}]$ and variance $[\sigma_x^2, \sigma_y^2]$, each calculated using the regions function from \S \ref{subsec:OutlierRejection};  map of the environment
\State Compute optimal policy for object, according to \S \ref{subsec:objectpolicy}
\While{$\mathbf{b}$ is not in goal region}
\State $\sigma^2 \gets \max{(\sigma_x,\sigma_y)}$
\If {$\sigma^2 > \sigma_{max}^2$}
\While{$\sigma^2 > \sigma_{min}^2$}
\State $ [x_{goal}, y_{goal}] \gets $ nearest corner in region
\State Apply \eqref{eq:PDcontrolPosition} to move toward $[x_{goal}, y_{goal}]$
\EndWhile
\Else  
%\If {$\mathrm{distance}( \mathbf{b}, [x_{goal}, y_{goal}] ) > k_1 r_b$}
%	\State$r_p \gets k_2 r_b$  \Comment{guarded move}
%	\Else
%	\State$r_p \gets k_3 r_b$  \Comment{pushing move}
%	\EndIf
%\State$r_p \gets k_3 r_b$  \Comment{pushing move}
\State Calculate $\mathbf{D}(\mathbf{b})$  \Comment{direction for object at $\mathbf{b}$}
\State Apply \eqref{eq:potentialfield}   \Comment{potential field for swarm}
\EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}
\centering
%\renewcommand{\figwid}{0.19\columnwidth}
%\href{http://youtu.be/tCej-9e6-4o}{\begin{overpic}[width =\figwid]{story1.png}\put(6,15){T = 5 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story2.png}\put(6,15){T = 12 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story3.png}\put(6,15){T = 20 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story4.png}\put(6,15){T = 25 s}
%\end{overpic}
%\begin{overpic}[width =\figwid]{story5.png}\put(6,15){T = 33 s}
%\end{overpic}}
\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}
\end{overpic}
\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 robots under automatic control.  See video in multimedia extension.
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure*}
\centering
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}

### 2016-08-16 journalSwarmControl/realExperiment.tex
Changed format

\begin{figure*}
\begin{center}
	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
	\begin{overpic}[width=0.5\columnwidth]{SetUp.pdf}%\put(1,75){A}
	\end{overpic}\begin{overpic}[width=0.25\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
	\end{overpic}\begin{overpic}[width=0.25\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
	\end{overpic}
\end{center}
\caption{\label{fig:setup}
Hardware platform. At right are the shapes used for hardware experiments and a visualization of the potential field. }
\end{figure*}

The objects were 3D printed from ABS plastic with a paper overlay. 
Shapes included a 325 cm$^2$ equilateral triangle, 
%Small Equilateral Triangle 140 cm$^2$,
 324 cm$^2$ square,
 281 cm$^2$ hexagon,
254 cm$^2$ circle, 
and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}.The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.

@@

\begin{figure*}
\begin{center}
	\begin{overpic}[width=\columnwidth]{RealMean}\put(42,20){\emph{t}= 90 s}\put(72,20){\emph{t}= 175 s}\end{overpic}
%	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
\end{center}
\vspace{-1em}
\caption{\label{fig:realMean}
Regulating average $x$ position of 100 kilobots using control law \eqref{eq:PDcontrolPosition}.
%Mean control plot with kilobots. %\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red circle in left hand corner of these, and put a scale bar in image}
}
\end{figure*}

@@

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,29){\emph{t} = 3 s}\put(38,29){\emph{t} = 410 s}\put(70,29){\emph{t} = 710 s}
%\put(15,22){\emph{t} = 1374 s}\put(49,22){\emph{t} = 2185 s}\put(83,22){\emph{t} = 2703 s}
%\end{overpic}
\begin{overpic}[width=.9\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
\put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control. The automatic controller classifies pink objects as obstacles and generates a policy to the goal. See  video in multimedia extension.}
%\vspace{-2em}
}
\end{figure}

>>>>

\begin{figure*}
\renewcommand{\figwid}{6cm}
\begin{center}
	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
	\begin{overpic}[height=\figwid]{SetUp.pdf}%\put(1,75){A}
	\end{overpic}\begin{overpic}[height=\figwid]{AllShapesR.pdf}%\put(-0,85){B}
	\end{overpic}\begin{overpic}[height=\figwid]{PotentialExp.pdf}%\put(-0,85){C}
	\end{overpic}
\end{center}
\caption{\label{fig:setup}
Hardware platform. At right are the shapes used for hardware experiments and a visualization of the potential field. }
\end{figure*}

The objects were 3D printed from ABS plastic with a paper overlay. 
Shapes included a 325 cm$^2$ equilateral triangle, 
%Small Equilateral Triangle 140 cm$^2$,
 324 cm$^2$ square,
 281 cm$^2$ hexagon,
254 cm$^2$ circle, 
and a 486 cm$^2$ (18 cm$\times$27 cm)  rectangle, all shown in Fig.~\ref{fig:setup}. The laser-cut patterns for the neon green fiducial markers on the robots and 3D files for objects are available at our github repository,~\cite{Shahrokhi2015GitHubShapeControl}.

@@

\begin{figure}
\begin{center}
	\begin{overpic}[width=1.0\columnwidth]{RealMean}\put(42,20){\emph{t}= 90 s}\put(72,20){\emph{t}= 175 s}\end{overpic}
%	\begin{overpic}[width=0.35\columnwidth]{XYMeanControl.eps}\end{overpic}
\end{center}
\vspace{-1em}
\caption{\label{fig:realMean}
Regulating mean $x$ position of 100 kilobots using control law \eqref{eq:PDcontrolPosition}.
%Mean control plot with kilobots. %\todo{place a red star and a red circle at two spots on graph, put two images of the swar at right with a red star and red circle in left hand corner of these, and put a scale bar in image}
}
\end{figure}

@@

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,29){\emph{t} = 3 s}\put(38,29){\emph{t} = 410 s}\put(70,29){\emph{t} = 710 s}
%\put(15,22){\emph{t} = 1374 s}\put(49,22){\emph{t} = 2185 s}\put(83,22){\emph{t} = 2703 s}
%\end{overpic}
\begin{overpic}[width=\columnwidth]{Snapshots.pdf}\put(6,80){\emph{t} = 3 s}\put(38,80){\emph{t} = 410 s}\put(70,80){\emph{t} = 710 s}
\put(15,8){\emph{t} = 1374 s}\put(49,8){\emph{t} = 2185 s}\put(83,8){\emph{t} = 2703 s}
\end{overpic}
\vspace{-1em}
\caption{\label{fig:expSnapShot}{Snapshots showing the object manipulation experiment with 100 kilobots under automatic control. The automatic controller classifies pink objects as obstacles and generates a policy to the goal. See  video in multimedia extension.}
%\vspace{-2em}
}
\end{figure}

### 2016-08-16 journalSwarmControl/theory.tex
added Mable's edits
Javier's comments fixed
Changed format

\subsection{Independent control of many robots is impossible}
A single robot is fully controllable. For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position of many robots}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} 
+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

We again analyze the controllability matrix $\mathcal{C}_{\mu}$:
\begin{equation}
\mathcal{C}_\mu=\left[ \begin{matrix} 
0\\
1
\end{matrix}
\,\middle\vert\,
 \begin{matrix} 
1\\
0
\end{matrix}
 \right],  \quad \textrm{rank}(\mathcal{C}_\mu)=2.
\end{equation}
Thus the average position and average velocity are controllable.

@@

We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}) \nonumber\\
u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}})  \label{eq:PDcontrolPosition}
\end{align}
here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

@@

\begin{algorithm}
\caption{Hybrid mean and variance control}\label{alg:MeanVarianceControl}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$  \Comment{initially not in variance control mode}
\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
\Loop
%\State  Compute $\sigma_x^2, \sigma_y^2$

@@

\begin{figure}
\centering
\begin{overpic}[width = 0.5\columnwidth]{hysteresis.pdf}
\put(33,40){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
\put(33,15){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
\begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:hysteresis} (Left) Two states to control mean and variance of a robot swarm. 
(Right) Switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The plot uses robot radius $r=1/10$.
%\vspace{-2em}
}
\end{figure}

@@

A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values shown in Fig.~\ref{fig:hysteresis}:
\begin{align}
 \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r) \nonumber\\
  \sigma^2_{max} &= 15r+ \sigma^2_{optimal}(n,r).
  \end{align}
%\sigma \approx 0.371258 n r,   \sigma_{min} = n*r*1/2, \sigma_{min} = n*r*3/2, 
%  this result was verified in http://www-math.mit.edu/~tchow/penny.pdf

>>>>

\subsection{Independent control of many robots is impossible}
A single robot is fully controllable. For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities. Without loss of generality, assume $m=1$.
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position of many robots}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, we create the following reduced order system that represents the mean $x$ position and velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} \\
&+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

We again analyze the controllability matrix $\mathcal{C}_{\mu}$:
\begin{equation}
\mathcal{C}_\mu=\left[ \begin{matrix} 
0\\
1
\end{matrix}
\,\middle\vert\,
 \begin{matrix} 
1\\
0
\end{matrix}
 \right],  \quad \textrm{rank}(\mathcal{C}_\mu)=2.
\end{equation}
Thus the mean position and mean velocity are controllable.

@@

We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}) \nonumber\\
u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}})  \label{eq:PDcontrolPosition}
\end{align}
 $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

@@

\begin{algorithm}
\caption{Hybrid mean and variance control}\label{alg:MeanVarianceControl}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ % \Comment{initially not in variance control mode}
\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
\Loop
%\State  Compute $\sigma_x^2, \sigma_y^2$

@@

\begin{figure}
\centering
\begin{overpic}[width = 0.8\columnwidth]{hysteresis.pdf}
\put(33,40){$\sigma^2(\mathbf{x}) < \sigma^2_{min}$ }
\put(33,15){$\sigma^2(\mathbf{x}) > \sigma^2_{max}$}\end{overpic}
%\begin{overpic}[width = 0.45\columnwidth]{VarianceMinMaxBand2v2.pdf}\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:hysteresis}  Hysteresis to control swarm mean and variance. 
%(Right) Switching conditions for variance control are set as a function of $n$, and designed to be larger than the optimal packing density. The plot uses robot radius $r=1/10$.
%\vspace{-2em}
}
\end{figure}

@@

A key challenge is to select proper values for $\sigma_{min}^2$ and $\sigma_{max}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values: 
%shown in Fig.~\ref{fig:hysteresis}:
\begin{align} \label{eq:sigMaxMin}
 \sigma^2_{min} &= 2.5r+ \sigma^2_{optimal}(n,r) \nonumber\\
  \sigma^2_{max} &= 15r+ \sigma^2_{optimal}(n,r).
  \end{align}
%\sigma \approx 0.371258 n r,   \sigma_{min} = n*r*1/2, \sigma_{min} = n*r*3/2, 
%  this result was verified in http://www-math.mit.edu/~tchow/penny.pdf
